POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types

### 2023-10-14 challenges/linearity.tex
updated some proofs

\begin{lemma}[Weakening]\mbox{}
  \label{lemma:weak}
  \begin{enumerate}
  \item If \( \types{\Gamma; \Delta}{P} \) then
    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
      \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
  \end{enumerate}
\end{lemma}
\begin{proof}
  By induction on the given derivations.
\end{proof}
\begin{lemma}[Strengthening]\mbox{}
  \label{lemma:strenD}
  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
\end{lemma}
\begin{proof}
  By induction on $P$, for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to the uniquely applicable rule.
  \begin{description}
    \item[$P = \Pend$] By inversion on the typing we have $\tend{\Delta}$. Apply \TirName{T-Inact}.
    \item[$P = \Ppar{P_0}{P_1}$] By inversion on the typing we have  $ \Delta, x : T = \Delta_0 , \Delta_1$.
      By cases on which environment $x$ is in, apply IH on that environment. 
    \item[$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By inversion, $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From the assumptions present we can apply IH.
  \end{description}
  The remaining cases are analogous.
\end{proof}

@@

\begin{theorem}[Type safety]
  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
\end{theorem}
\begin{proof}
  In order to prove that $P$ is well-formed, let us consider any
  process of the form
  $\Presd{x_1}{y_1}{\dots
    \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}}$ that is
  structurally congruent to $P$. Moreover, assume that $P_1$ is
  prefixed at some $x_1$ and $P_2$ is prefixed at some $x_2$ such that
  $P_1\equiv\Pout{x}{v}{P_1'}$ (the opposite case proceeds
  similarly). We need to show that $P_2\equiv\Pin{x}{v}{P_2'}$. This
  can be easily done by contradiction. In fact, if
  $P_2\equiv\Pout{x}{v}{P_2'}$ then the typing rule for restriction
  would be violated since the type of $x_1$ and $y_1$ cannot be dual.
  % 
  % We proceed by induction on the definition of $\types{}{}$. The key
  % cases are that for parallel composition and that for name
  % restriction. In the case of parallel composition, linearity (and
  % context splitting) ensure that every process in parallel will have
  % different channel names. Then, the restriction rule enforces
  % duality, therefore an output will be matched by an input.
\end{proof}

>>>>

\begin{lemma}[Weakening]\mbox{}
  \label{lemma:weak}
  \begin{enumerate}
  \item If \( \types{\Gamma; \Delta}{P} \) then
    \( \types{(\Cadd{\Gamma}{l});\Delta}{P} \).
      \item If \( \types{\Gamma; \Delta}{P} \) and \( \tend{\Delta} \) then
    \( \types{\Gamma;(\Cadd{\Delta}{\hastype{x}{\Tend}})}{P} \).
  \end{enumerate}
\end{lemma}
\begin{proof}
  By induction on the given derivations.
\end{proof}
\begin{lemma}[Strengthening]\mbox{}
  \label{lemma:strenD}
  If \( \types{\Gamma; (\Delta, x : T)}{P} \) and
  $x\not\in \freenames{P}$, then \( \types{\Gamma; \Delta}{P} \).
\end{lemma}
\begin{proof}
  By induction on the last applied rule of the proof of
  $\types{\Gamma; (\Delta, x : T)}{P}$.
  % , for each case inverting $\types{\Gamma; (\Delta, x : T)}{P}$ to
  % the uniquely applicable rule.
  \begin{description}
  \item[\TirName{T-End}] Since $\tend{\Delta}$, we can just reapply
    the rule without $x:T$.
    % [$P = \Pend$] we Apply \TirName{T-Inact}.

  \item[\TirName{T-Par}] In this case, we have that
    $ \Delta, x : T = \Delta_0 , \Delta_1$.  By cases on which
    environment $x$ is in, we just apply the induction hypothesis on
    that environment.
    %
    %
    % [$P = \Ppar{P_0}{P_1}$] By inversion on the typing we have  $ \Delta, x : T = \Delta_0 , \Delta_1$.
    %   By cases on which environment $x$ is in, apply IH on that environment. 
  \item[\TirName{T-Res}] Without loss of generality, we assume that
    $x \notin \{y,z\}$. Since
    $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
    the assumptions present we can apply the induction hypothesis.
    % inversion,
    % $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
    % the assumptions present we can apply IH.
    %
    % [$P = \Presd{y}{z}{P}$] We may assume $x \notin \{y,z\}$. By
    % inversion,
    % $\Delta, x : T = \Delta', y : T_0, z : \dual{T_1}, x : T $. From
    % the assumptions present we can apply IH.
  \end{description}
  The remaining cases are analogous.
\end{proof}

@@

\begin{theorem}[Type safety]
  If \( \types{\Cempty}{P} \), then \( P \) is well-formed.
\end{theorem}
\begin{proof}
  In order to prove that $P$ is well-formed, let us consider any
  process of the form
  $\Presd{x_1}{y_1}{\dots
    \Presd{x_n}{y_n}{(\Ppar{\Ppar{P_1}{P_2}}{R})}}$ that is
  structurally congruent to $P$. Clearly, by Lemma~\ref{le:presequiv},
  well-typedness must be preserved by structural congruence. Moreover,
  assume that $P_1$ is prefixed at $x_1$ and $P_2$ is prefixed at
  $y_1$ such that $P_1\equiv\Pout{x_1}{v}{P_1'}$ (the opposite case
  proceeds similarly). We need to show that
  $P_2\equiv\Pin{y_1}{v}{P_2'}$. This can be easily done by
  contradiction. In fact, if $P_2\equiv\Pout{x}{v}{P_2'}$ then the
  typing rule for restriction would be violated since the type of
  $x_1$ and $y_1$ cannot be dual.
  % 
  % We proceed by induction on the definition of $\types{}{}$. The key
  % cases are that for parallel composition and that for name
  % restriction. In the case of parallel composition, linearity (and
  % context splitting) ensure that every process in parallel will have
  % different channel names. Then, the restriction rule enforces
  % duality, therefore an output will be matched by an input.
\end{proof}
